---
title: "R Cheatsheet"
author: "Patrice Robin, Patrik Widmer"
date: "4 2 2021"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Shortcuts

There are several useful shortcuts in R. And as Benjamin Frankling allegedly said: "time is money":

-   Control + Shift + C --\> Comment out a marked chunk
-   Control + Shift + A --\> Auto format code
-   Control + Shift + S --\> Source where your are working in
-   Control + Alt + I --\> New R Code part in R Markdown

## Clean Data

### Janitor

You need a nice solution to clean your data, the janitor is here to clean the mess:

-   <https://rstudio-conf-2020.github.io/r-for-excel/tidying.html>

With Janitor we can easily adjust columns names, see all rows that are duplicates within variables you choose and remove empty columns or rows.

clean_names produces by default **lower_snake_case**. However other formats like **snake**, **lower_camel**, **lower_upper** and **upper_lower** is also possible

```{r janitor, eval=FALSE, echo=true}
#install.packages("janitor")
library("janitor")
janitor::clean_names()
janitor::get_dupes()
janitor::remove_empty()
```

```{r}

```
https://www.onceupondata.com/2019/01/25/ggplot2-divergent-bars/
