---
title: "Impact of Covid on Cyclists and Pedestrians in Zurich"
author: "Patrice.Robin, Patrik.Widmer"
date: "1 2 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE ,
                      message=FALSE,
                      warning = FALSE,
                      results = "hide"
                      )

```

# Introduction

Covid19 had a huge impact on our society. It changed the way we interact on a professional and social level. Home offices, lockdowns, and partially closed stores have had a particularly severe impact on urban centers.

In this paper, we will examine the impact of the pandemic on the number of pedestrians and bicycles in different areas of Zurich.

The city of Zurich automatically collects data on pedestrians and bicyclists through counting loops installed in the ground in different areas of the city. We can get a clear view of the impact of Covid19 on our daily lives by looking at the data of pedestrians and cyclists in 2020 and comparing it to the previous year.

```{r load packages}
# this line of code only runs, if these packages are required
if (!require(tidyverse)) {install.packages('"tidyverse"')}
if (!require(ggplot2)) {install.packages('ggplot2')}
if (!require(scales)) {install.packages('scales')}
if (!require(rgdal)) {install.packages('rgdal')}
if (!require(sf)) {install.packages('sf')}
if (!require(rasterVis)) {install.packages('rasterVis')}
if (!require(dplyr)) {install.packages('dplyr')}
if (!require(av)) {install.packages('av')}
if (!require(ggpubr)) {install.packages('ggpubr')}
```

```{r english_local}
#set local working directory to English
Sys.setlocale("LC_TIME", "English")
```

# Data

All the data regarding Zurich are pulled from the official website of the city of Zurich *Open Data Stadt Zürich*. The fallowing information are used in this paper:

-   Number of pedestrians and bicycles counted in 2020 and 2019

-   The location of the counting stations.

-   The different *Stadtkreise* and their geodata points.

-   The different districts in Zurich with the corresponding geo data points.

-   A categorization of different areas based on their usage and building development.

The Covid19 data for Switzerland including the categories *cases, deaths, test, hospitalization, and vaccinations* are based on the research of Roser et al. (2020) and pulled from the *Our World in Data* website. In this paper, we will use the smoothed numbers of newly infected people in Switzerland. We decided to use the smoothed numbers since the newly reported cases are directly linked to the working week of medical staff, laboratories, and the Federal Office of Public Health Switzerland (FOPH). As a result, often fewer cases are reported on the weekend and more cases are listed in the beginning of the working week.

Due to the smoothing of the data, the first recorded Covid19 case in our dataset is from 01.03.2020, while the first recorded case actually appeared on the 25.02.2020.

```{r load_covid}
#load covid data
#source : https://ourworldindata.org/coronavirus/country/switzerland?country=~CHE
covid <- read.csv("covid-data.csv", encoding="UTF-8")
covid$date <- as.Date(covid$date,format = "%d.%m.%y")

```

```{r load_zwalks}
#load pedestrian and cyclist data
#source: https://data.stadt-zuerich.ch/dataset/ted_taz_verkehrszaehlungen_werte_fussgaenger_velo
zwalks2020 <- read.csv("2020_verkehrszaehlungen_werte_fussgaenger_velo.csv") #2019
zwalks2019 <- read.csv("2019_verkehrszaehlungen_werte_fussgaenger_velo.csv") #2020

zwalks <- dplyr::union(zwalks2020, zwalks2019)

```

```{r load_kreis}
#load Stadtkreise
#source https://data.stadt-zuerich.ch/dataset/geo_stadtkreise
zkreis <- st_read("./Stadtkreise/stzh.adm_stadtkreise_v.shp")

```

```{r load_zstation}
#load countingstations
#source https://data.stadt-zuerich.ch/dataset/geo_standorte_der_automatischen_fuss__und_vlozaehlungen

zstation <- st_read("./Countingstation/taz.view_eco_standorte.shp")
```

```{r load_zdistricts}
#Districts
#source: https://data.stadt-zuerich.ch/dataset/ktzh_quartieranalyse
zdistricts <- st_read("./Quartieranalyse/QUARTIERE_F.shp")
```

```{r load_zzones}
#Bauzonen
#sources https://data.stadt-zuerich.ch/dataset/geo_nutzungsplanung___kommunale_bau__und_zonenordnung__bzo__jahresendstand_2020
zzones <- st_read("./Bauzonen/afs.bzo_zone_v.shp")
```

## Preparation

In the initial dataset summarizes the counted pedestrians and cyclists into time frames of 15 minutes. Due to our interest into daily, weekly and monthly changes, we bring the counting data into needed format. Additionally, several adjustment to the data are made for a simpler comparison. The counting stations indicate whether an individual is going into or going out of a center. Since we are interested in the total number of individuals walking or cycling in the area, we summarize them into one number each:

-   Total cyclists
-   Total Pedestrians

Furthermore, we also summarize those two numbers into one single number containing all the measured pedestrians and cyclists at each location:

-   Total People

```{r clean_data}
#extract date from datetime
zwalks['date'] <- as.Date(zwalks$DATUM)

#lower cases column names
zwalks <- setNames(zwalks, tolower(names(zwalks)))

#replace NA with 0
zwalks[is.na(zwalks)] <- 0

#rename columns
zwalks <- zwalks %>%
  rename(datetime = datum)
```

```{r summarize_daily}
#summarize data on a daily basis
zwalks.day <-
  aggregate(zwalks[c("velo_in", "velo_out", "fuss_in", "fuss_out")],
            by = zwalks[c("fk_zaehler", "fk_standort", "date", "ost", "nord")],
            FUN = sum)

#summarize bicycles and pedestrians
zwalks.day['fuss_all'] <- zwalks.day$fuss_in + zwalks.day$fuss_out
zwalks.day['velo_all'] <- zwalks.day$velo_in + zwalks.day$velo_out
zwalks.day['people_all'] <-
  zwalks.day$velo_all + zwalks.day$fuss_all
```

In a second step, we join the smoothed Covid data to our daily records of cyclists and pedestrians. Before 01.03.2020 there are no Covid cases.

```{r zwalks_date}
#add months and weekdays
zwalks.day$year <- format(zwalks.day$date, "%Y")
zwalks.day$month <- format(zwalks.day$date, "%b")
zwalks.day$week <- format(zwalks.day$date, "%W") #create weeks from date
zwalks.day$week <- as.numeric(zwalks.day$week) + 1 #make numeric and add 1 so that it goes from 1 to 53

#weekdays
zwalks.day$weekday <- weekdays(zwalks.day$date)

weekday_order <- c( "Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday","Sunday")
```

```{r join_zwalks_covid}
# overall pedestrians and bicycles
zwalks.agg <-
  aggregate(zwalks.day[c("people_all", "fuss_all", "velo_all")],
            by = zwalks.day[c( "date", "month", "weekday", "week","year")],
            FUN = sum)

# left join zwalks agg and covid on date
zwalks.covid <- zwalks.agg %>%
  left_join(dplyr::select(covid,
                          date,
                          new_cases_smoothed),
            by = "date")

```

```{r agg_monthly}
#replace NA with 0 in covid cases
zwalks.covid[is.na(zwalks.covid)] <- 0

zwalks.covid.month <-
  aggregate(zwalks.covid[c("people_all", "fuss_all", "velo_all","new_cases_smoothed")],
            by = zwalks.covid[c( "year", "month")],
            FUN = sum)



#pivot wide table on months
zwalks.month.wide <- zwalks.covid.month %>%
  dplyr::select(year, people_all, month) %>%
  pivot_wider(names_from = year,
              values_from = people_all)
```

Based on the information of the Swiss Tourism Federation (2021) we additionally the two lockdowns in 2020. The first lockdown from March 16th 2020 until Mai 11th 2020. The second lockdown starting in 22nd December 2020 and ongoing in the next year.

```{r}
### define lockdown period
### chronic of lockdowns https://www.stv-fst.ch/de/chronik-coronavirus
lockdown.1 <- seq(as.Date("2020/03/16"), by = "day", length.out = 57)
lockdown.2 <- seq(as.Date("2020/12/22"), by = "day", length.out = 10)

### add lockdown flag to dataset zwalks.covid
zwalks.covid$lockdown <- FALSE
zwalks.covid$lockdown01 <- 0

zwalks.covid[zwalks.covid$date %in% lockdown.1 , ]$lockdown <- TRUE
zwalks.covid[zwalks.covid$date %in% lockdown.1 , ]$lockdown01 <- 1

zwalks.covid[zwalks.covid$date %in% lockdown.2 , ]$lockdown <- TRUE
zwalks.covid[zwalks.covid$date %in% lockdown.1 , ]$lockdown01 <- 1

```

# Visualization

```{r month_plots, results="markup"}
#Aggregated Data - Daily 2019/2020
p.months1 <- ggplot(zwalks.covid, aes(x=date, y=people_all)) +
  geom_point(na.rm = TRUE, size = 0.75) +
  xlab("") +
  ylab("Daily per Location ") +
  scale_x_date(
    labels = date_format("%b"),
    breaks = "1 month") + 
  theme(axis.text.x = element_text(angle = 45)) +
  geom_point(aes(colour = factor(year)), size = 2)+
  theme(legend.position = "none")

# 
# p.months2 <- ggplot(zwalks.covid, aes(x=factor(month, level=month.abb), y=people_all)) +
#   geom_point(na.rm = TRUE, size = 0.75) +
#   xlab(" ") +
#   ylab("Total Pedestrians and Cyclists") +
#   theme(axis.text.x = element_text(angle = 45)) +
#   geom_point(aes(colour = factor(year)), size = 4)


p.months3 <- ggplot(zwalks.covid.month, aes(x=factor(month, level=month.abb), y=people_all))+
  geom_line(na.rm = TRUE, size = 0.75) +
  xlab("") +
  ylab("Monthly Total") +
  theme(axis.text.x = element_text(angle = 45)) +
  geom_point(aes(colour = factor(year)), size = 2)+
  labs(color = "Year") 


p.months4 <-ggplot(zwalks.covid.month, aes(x=factor(month, level=month.abb), y=people_all, group=year, colour=year))+
  geom_line(size=2)+
  xlab("") +
  ylab(" ") +
  theme(axis.text.x = element_text(angle = 45))+
  scale_fill_discrete(name = "Year")

plot.months <- ggarrange(
  p.months1,
  ggarrange(
    p.months3,
    p.months4,
    common.legend = TRUE,
    legend = "bottom"
  ),
  nrow = 2
)

plot.months <- annotate_figure(
  plot.months,
  top = text_grob(
    "Pedestirans and cyclists in 2019 and 2020",
    color = "black",
    face = "bold",
    size = 14),
    left = text_grob("Counted Pedestrians and Bicycles",vjust=-.02, color = "black", rot = 90, hjust=0.5)
  )

plot.months
```

# Covid and Lockdown

In the next step we take a look ath how the smoothed covid cases and the counted pedestrians and cyclists are plotted against each other. We only include the counting data from March 1st 2020 until December 31st 2020.

By adding a regression line, we can see that there is a negative correlation between the number of cases and the number of counted people. However, the covnfidence intervall of the trend line is not very accurate and an inexisting confidenceintervall with low covid cases.

The clear right scewing of the dat is not a surpise, as we had most of the time below 1000 new covid cases per day. The big variety of recorded people in the street are likely due to whether, temperatures, seasonality but also due to the lockdown. The lockdwon is a political decision on not directly triggered or linked to the number of recorded cases.


Additionaly, we take a look at the counted people during the two lockdowns and non-lockdown period. The lockdown boxplot clearly indicates that there is a difference between the counted people during lockdown and during non-lockdown, with fewer people on the street during the two periods of lockdown.

In a third look at the relationship between the recorded cyclists and pedestrians and the two lockdowns by applying a logistic regression. It becomes clear, that the lockdown has indeed a negative impact on the number of recorded people on the street.
```{r covid_plots, results="markup"}
###  Boxplot
covid.box <-
  ggplot(
    data = zwalks.covid %>% filter(date >= "2020-03-01"),
    mapping = aes(y = people_all,
                  x = lockdown)
  ) +
  geom_boxplot() +
  geom_hline(yintercept = 0) +
  ylab("People") +
  xlab("Lockdown") +
  ylim(0, 125000)

###  Plot People vs Cases
covid.plot <-
  ggplot(data = zwalks.covid %>% filter(date >= "2020-03-01"),
         aes(y = people_all,
             x = new_cases_smoothed)) +
  geom_point() +
  xlab("New Covid Cases Smoothed") +
  ylab("People") +
  ylim(0, 125000) +
  geom_smooth(method = "lm")

### Logistic regression
zwalks.covid.glm <-
  glm(lockdown ~  people_all, family = "binomial",
      data = zwalks.covid %>% filter(date >= "2020-03-01"))

summary(zwalks.covid.glm)


### Plot the Logistic regression
covid.lr1 <- ggplot(data = zwalks.covid %>% filter(date >= "2020-03-01"),
                   mapping = aes(y = lockdown01,
                                 x = people_all)) +
  xlab("Pedestrians and Cyclists") +
  ylab("Lockdown")+
  geom_point() +
  geom_smooth(
    method = "glm",
    se = FALSE,
    method.args = list(family = "binomial")
  )



p.covid <- ggarrange(covid.plot,
                     covid.box,
                     covid.lr1,
                     align = "hv",
                     nrow = 3)


p.covid <- annotate_figure(
  p.covid,
  top = text_grob(
    "Pedestrians and cyclists in 2020 compared to covid cases and lockdown",
    color = "black",
    face = "bold",
    size = 14
  )
)

p.covid

```


# Conclusion

It would be interesting to quantify the impact of the two lockdown on the number of pedestrians and cyclists statistically. For this we a longer time series than just two years.
