---
title: "Impact of Covid on Cyclists and Pedestrians in Zurich"
author: "Patrice Robin, Patrik Widmer"
date: "25.05.2021"
output:
  html_document:
    toc: true # table of content true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: true  ## if you want number sections at each table
bibliography: "./Sources/packages.bib"
nocite: '@*'
---

<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->

\tableofcontents

<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->

<!-- Setting global options (hidden section) -->

<!-- Load packages (hidden section) -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE ,
                      message=FALSE,
                      warning = FALSE,
                      results = "hide"
                      )

```

```{r load packages}
# -------------------------- Load Packages --------------------------
# this line of code only runs, if these packages are required
if (!require(tidyverse)) {install.packages('"tidyverse"')}
if (!require(ggplot2)) {install.packages('ggplot2')}
if (!require(scales)) {install.packages('scales')}
if (!require(rgdal)) {install.packages('rgdal')}
if (!require(sf)) {install.packages('sf')}
if (!require(rasterVis)) {install.packages('rasterVis')}
if (!require(dplyr)) {install.packages('dplyr')}
if (!require(av)) {install.packages('av')}
if (!require(ggpubr)) {install.packages('ggpubr')}
if (!require(gdtools)) {install.packages('gdtools')}
if (!require(kableExtra)) {install.packages("kableExtra", dependencies = TRUE)}
if (!require(janitor)) {install.packages("janitor")}
if (!require(stargazer)) {install.packages("stargazer")}
if (!require(vembedr)) {install.packages("vembedr")}
if (!require(rmarkdown)) {install.packages("rmarkdown")}
if (!require(devtools)) {install.packages("devtools")}
if (!require(usethis)) {install.packages("usethis")}

```

```{r english_local}
# set local working directory to English
Sys.setlocale("LC_TIME", "English")
```

<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->

# Introduction

Covid19 had a huge impact on our society. It changed the way we interact on a professional and social level. Home offices, lockdowns, and partially closed stores have had a particularly severe impact on urban centers.

In this paper, we will examine the impact of the pandemic on the number of pedestrians and bicycles in different areas of Zurich.

The city of Zurich automatically collects data on pedestrians and bicyclists through counting loops installed in the ground in different areas of the city. We can get a clear view of the impact of Covid19 on our daily lives by looking at the data of pedestrians and cyclists in 2020 and comparing it to the previous year.

<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->

<!-- Load data -->

# Data

All the data regarding Zurich are pulled from the official website of the city of Zurich *Open Data Stadt Zürich* (2021). The fallowing information are used in this paper:

-   Number of pedestrians and bicycles counted in 2020 and 2019

-   The location of the counting stations.

-   The different *Stadtkreise* and their geodata points.

-   The different districts in Zurich with the corresponding geo data points.

-   A categorization of different areas based on their usage and building development.

The Covid19 data for Switzerland including the categories *cases, deaths, test, hospitalization, and vaccinations* are based on the research of Roser et al. (2020) and pulled from the *Our World in Data* website. In this paper, we will use the smoothed numbers of newly infected people in Switzerland. We decided to use the smoothed numbers since the newly reported cases are directly linked to the working week of medical staff, laboratories, and the Federal Office of Public Health Switzerland (FOPH). As a result, often fewer cases are reported on the weekend and more cases are listed in the beginning of the working week.

Due to the smoothing of the data, the first recorded Covid19 case in our dataset is from 01.03.2020, while the first recorded case actually appeared on the 25.02.2020.

```{r load_covid}
# load covid data
# source : https://ourworldindata.org/coronavirus/country/switzerland?country=~CHE
covid <- read.csv("./Data/covid-data.csv", encoding="UTF-8")
covid$date <- as.Date(covid$date,format = "%d.%m.%y")

```

```{r load_zwalks}
# load pedestrian and cyclist data
# source: https://data.stadt-zuerich.ch/dataset/ted_taz_verkehrszaehlungen_werte_fussgaenger_velo
zwalks2020 <- read.csv("./Data/2020_verkehrszaehlungen_werte_fussgaenger_velo.csv") # 2019
zwalks2019 <- read.csv("./Data/2019_verkehrszaehlungen_werte_fussgaenger_velo.csv") # 2020

zwalks <- dplyr::union(zwalks2020, zwalks2019)

```

```{r load_kreis}
# load Stadtkreise
# source https://data.stadt-zuerich.ch/dataset/geo_stadtkreise
zkreis <- st_read("./Data/Stadtkreise/stzh.adm_stadtkreise_v.shp")

```

```{r load_zstation}
# load countingstations
# source https://data.stadt-zuerich.ch/dataset/geo_standorte_der_automatischen_fuss__und_vlozaehlungen

zstation <- st_read("./Data/Countingstation/taz.view_eco_standorte.shp")
zstation <- cbind(zstation, st_coordinates(zstation)) # split coordinates into X and Y

```

```{r load_zdistricts}
# Districts
# source: https://data.stadt-zuerich.ch/dataset/ktzh_quartieranalyse
zdistricts <- st_read("./Data/Quartieranalyse/QUARTIERE_F.shp")

```

```{r load_zzones}
# Bauzonen
# sources https://data.stadt-zuerich.ch/dataset/geo_nutzungsplanung___kommunale_bau__und_zonenordnung__bzo__jahresendstand_2020
zzones <- st_read("./Data/Bauzonen/afs.bzo_zone_v.shp")
```

## Preparation

In the initial dataset summarizes the counted pedestrians and cyclists into time frames of 15 minutes. Due to our interest into daily, weekly and monthly changes, we bring the counting data into needed format. Additionally, several adjustment to the data are made for a simpler comparison. The counting stations indicate whether an individual is going into or going out of a center. Since we are interested in the total number of individuals walking or cycling in the area, we summarize them into one number each:

-   Total cyclists
-   Total Pedestrians

Furthermore, we also summarize those two numbers into one single number containing all the measured pedestrians and cyclists at each location:

-   Total People

```{r clean_data}
# extract date from datetime
zwalks['date'] <- as.Date(zwalks$DATUM)

# lower cases column names
zwalks <- setNames(zwalks, tolower(names(zwalks)))

# replace NA with 0
zwalks[is.na(zwalks)] <- 0

# rename columns
zwalks <- zwalks %>%
  rename(datetime = datum)
```

```{r summarize_daily}
# -------------------------- Aggregate People --------------------------
# summarize data on a daily basis
zwalks.day <-
  aggregate(zwalks[c("velo_in", "velo_out", "fuss_in", "fuss_out")],
            by = zwalks[c("fk_zaehler", "fk_standort", "date", "ost", "nord")],
            FUN = sum)

# summarize bicycles and pedestrians
zwalks.day['fuss_all'] <- zwalks.day$fuss_in + zwalks.day$fuss_out
zwalks.day['velo_all'] <- zwalks.day$velo_in + zwalks.day$velo_out
zwalks.day['people_all'] <- zwalks.day$velo_all + zwalks.day$fuss_all
```

In a second step, we join the smoothed Covid data to our daily records of cyclists and pedestrians. Before 01.03.2020 there were no covid cases.

```{r zwalks_date}
# -------------------------- Date Format --------------------------
# add months and weekdays
zwalks.day$year <- format(zwalks.day$date, "%Y")
zwalks.day$month <- format(zwalks.day$date, "%b")
zwalks.day$week <- format(zwalks.day$date, "%W") # create weeks from date
zwalks.day$week <- as.numeric(zwalks.day$week) + 1 # make numeric and add 1 so that it goes from 1 to 53

# weekdays
zwalks.day$weekday <- weekdays(zwalks.day$date)

weekday_order <- c( "Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday","Sunday")
```

```{r join_zwalks_covid}
# overall pedestrians and bicycles
zwalks.agg <-
  aggregate(zwalks.day[c("people_all", "fuss_all", "velo_all")],
            by = zwalks.day[c( "date", "month", "weekday", "week","year")],
            FUN = sum)

# left join zwalks agg and covid on date
zwalks.covid <- zwalks.agg %>%
  left_join(dplyr::select(covid,
                          date,
                          new_cases_smoothed),
            by = "date")

```

```{r agg_monthly}
# replace NA with 0 in covid cases
zwalks.covid[is.na(zwalks.covid)] <- 0

zwalks.covid.month <-
  aggregate(zwalks.covid[c("people_all", "fuss_all", "velo_all","new_cases_smoothed")],
            by = zwalks.covid[c( "year", "month")],
            FUN = sum)


# pivot wide table on months
zwalks.month.wide <- zwalks.covid.month %>%
  dplyr::select(year, people_all, month) %>%
  pivot_wider(names_from = year,
              values_from = people_all)
```

Based on the information of the Swiss Tourism Federation (2021) we additionally the two lockdowns in 2020. The first lockdown from March 16th 2020 until Mai 11. 2020. The second lockdown starting in 22. December 2020 and ongoing in the next year.

```{r}
# -------------------------- Lockdown Periods --------------------------
### define lockdown period
### chronic of lockdowns https://www.stv-fst.ch/de/chronik-coronavirus
lockdown.1 <- seq(as.Date("2020/03/16"), by = "day", length.out = 57)
lockdown.2 <- seq(as.Date("2020/12/22"), by = "day", length.out = 10)

### add lockdown flag to dataset zwalks.covid
zwalks.covid$lockdown <- FALSE
zwalks.covid$lockdown01 <- 0

zwalks.covid[zwalks.covid$date %in% lockdown.1 , ]$lockdown <- TRUE
zwalks.covid[zwalks.covid$date %in% lockdown.1 , ]$lockdown01 <- 1

zwalks.covid[zwalks.covid$date %in% lockdown.2 , ]$lockdown <- TRUE
zwalks.covid[zwalks.covid$date %in% lockdown.1 , ]$lockdown01 <- 1

```

## Counting stations

In the graphic below the counting stations are visualized. The counting stations are not evenly distributed over the twelve districts. The table below shows that district 1 has the most counting stations with a total of 9 and district 7 has no counting stations. We believe that the city of Zurich deliberately selected locations that are highly frequented. Consequently, most counting locations are in the city center or in regional center of Zurich.

A total of 48 different counting location were used to collect the numbers of pedestriand and cyclists during 2019 and 2020.

```{r counting.stations}
# -------------------------- Mapping Counting Stations --------------------------

# plot Stadtkreise and counting stations
plot.map <- ggplot()+
  geom_sf(data = zkreis, size = 0, color = "black", fill="lightblue")+
  geom_sf(data = zzones, size = 0.25, color = "darkgrey", fill="white")

plot.kreis <- plot.map +
    geom_sf(data = zkreis, size = 1, color = "black", fill= NA )+
    geom_sf_text(data = zkreis, aes(label = knr), colour = "black", size=10)
  
# counting stations  
plot.station <- plot.kreis + geom_point(data = zstation, aes(x=X, y=Y), size = 3, 
                                    shape = 23, fill = "red")
  
plot.station


# join the shape files - left on station
zstation.kreis <- st_join(zstation, zkreis, 
                  join = st_within)


```

```{r table.stations, results="markup"}
# -------------------------- Counting Stations --------------------------

# left Stadtkreise
counting.stations <- tibble(District = 1:12) %>% # empty tibble with 12 district
  left_join(zstation.kreis %>% # left join
  group_by(knr) %>%
  summarise(length(unique(abkuerzung))) %>%
  rename("Quantity of Stations" = "length(unique(abkuerzung))"),
    by = c("District" = "knr"))

counting.stations <- counting.stations[0:2]

counting.stations <- counting.stations %>%
  adorn_totals("row") 


options(knitr.kable.NA = "0")

knitr::kable(counting.stations,
             align = "cr",
             caption = "Couting Stations per District") %>%
  kable_styling(bootstrap_options = "striped") %>%  # from the kabelExtra package
  row_spec(7, color = 'red', bold = TRUE) # District 7 has 0 counting stations --> mark red and bold


```

# Analysis

In the main chapter, we look at the counted pedestrians and cyclists counted in 2019 and 2020 within the city of Zurich. We look at the distribution over time, the difference between the 12 Stadtkreise, the difference between the zones the counting stations are positioned in, and lastly the impact of the two lockdowns.

## Months

First, we take a closer look at the counted pedestrians and cyclists counted together in 2020 and 2019. A comparison of the monthly data is presented below. The top graph below shows the summarized pedestrians and cyclists for each measurement location for the last two years. The two graphs below show the total of measured people for each month in 2019 and 2020 in a direct comparison.

It is obvious that there were more people on the streets in July, August and September in 2020 than in the previous year. We argue that an explanation are the continuous travel restrictions coming from and going to other countries combined with the eased lockdown within Switzerland.Hence, people spent their holidays at home or simply kept working.

```{r month_plots, results="markup"}
# -------------------------- Counted Pedestrians and Cyclists --------------------------

# Aggregated Data - Daily 2019/2020
p.months1 <- ggplot(zwalks.covid, aes(x=date, y=people_all)) +
  geom_point(na.rm = TRUE, size = 0.75) +
  xlab("") +
  ylab("Daily per Location ") +
  scale_x_date(
    labels = date_format("%b"),
    breaks = "1 month") + 
  theme(axis.text.x = element_text(angle = 45)) +
  geom_point(aes(colour = factor(year)), size = 2)+
  theme(legend.position = "none")

# 
# p.months2 <- ggplot(zwalks.covid, aes(x=factor(month, level=month.abb), y=people_all)) +
#   geom_point(na.rm = TRUE, size = 0.75) +
#   xlab(" ") +
#   ylab("Total Pedestrians and Cyclists") +
#   theme(axis.text.x = element_text(angle = 45)) +
#   geom_point(aes(colour = factor(year)), size = 4)


p.months3 <- ggplot(zwalks.covid.month, aes(x=factor(month, level=month.abb), y=people_all))+
  geom_line(na.rm = TRUE, size = 0.75) +
  xlab("") +
  ylab("Monthly Total") +
  theme(axis.text.x = element_text(angle = 45)) +
  geom_point(aes(colour = factor(year)), size = 2)+
  labs(color = "Year") 


p.months4 <-ggplot(zwalks.covid.month, aes(x=factor(month, level=month.abb), y=people_all, group=year, colour=year))+
  geom_line(size=2)+
  xlab("") +
  ylab(" ") +
  theme(axis.text.x = element_text(angle = 45))+
  scale_fill_discrete(name = "Year")

plot.months <- ggarrange(
  p.months1,
  ggarrange(
    p.months3,
    p.months4,
    common.legend = TRUE,
    legend = "bottom"
  ),
  nrow = 2
)

plot.months <- annotate_figure(
  plot.months,
  top = text_grob(
    "Pedestrians and cyclists in 2019 and 2020",
    color = "black",
    face = "bold",
    size = 14),
    left = text_grob("Counted Pedestrians and Bicycles", color = "black", rot = 90, hjust=0.5)
  )

plot.months
```

## Covid and Lockdown

In the next step we take a look at how the smoothed covid cases and the counted pedestrians and cyclists are plotted against each other. We only include the counting data from March 1st 2020 until December 31st 2020.

By adding a regression line, we can see that there is a negative correlation between the number of cases and the number of counted people. However, the confidence interval of the trend line is not very accurate and an in existing confidence interval with low covid cases.

The clear right skewing of the data is not a surprise, as we had most of the time below 1000 new covid cases per day. The big variety of recorded people in the street are likely due to whether, temperatures, seasonality but also due to the lockdown. The lockdown is a political decision on not directly triggered or linked to the number of recorded cases.

Additionally, we take a look at the counted people during the two lockdowns and non-lockdown period. The lockdown boxplot clearly indicates that there is a difference between the counted people during lockdown and during non-lockdown, with fewer people on the street during the two periods of lockdown.

In a third look at the relationship between the recorded cyclists and pedestrians and the two lockdowns by applying a logistic regression. It becomes clear, that the lockdown has indeed a negative impact on the number of recorded people on the street.

```{r covid_plots, results="markup"}
# -------------------------- Covid and Lockdown --------------------------

###  Boxplot
covid.box <-
  ggplot(
    data = zwalks.covid %>% filter(date >= "2020-03-01"),
    mapping = aes(y = people_all,
                  x = lockdown)
  ) +
  geom_boxplot() +
  geom_hline(yintercept = 0) +
  ylab("People") +
  xlab("Lockdown") +
  ylim(0, 125000)

###  Plot People vs Cases
covid.plot <-
  ggplot(data = zwalks.covid %>% filter(date >= "2020-03-01"),
         aes(y = people_all,
             x = new_cases_smoothed)) +
  geom_point() +
  xlab("New Covid Cases Smoothed") +
  ylab("People") +
  ylim(0, 125000) +
  geom_smooth(method = "lm")

### Logistic regression
zwalks.covid.glm <-
  glm(lockdown ~  people_all, family = "binomial",
      data = zwalks.covid %>% filter(date >= "2020-03-01"))


### Plot the Logistic regression
covid.lr1 <- ggplot(data = zwalks.covid %>% filter(date >= "2020-03-01"),
                   mapping = aes(y = lockdown01,
                                 x = people_all)) +
  xlab("Pedestrians and Cyclists") +
  ylab("Lockdown")+
  geom_point() +
  geom_smooth(
    method = "glm",
    se = FALSE,
    method.args = list(family = "binomial")
  )


# arrange plots
p.covid <- ggarrange(covid.plot,
                     covid.box,
                     covid.lr1,
                     align = "hv",
                     nrow = 3)


# annotate figures
p.covid <- annotate_figure(
  p.covid,
  top = text_grob(
    "Pedestrians and cyclists in 2020 compared to covid cases and lockdown",
    color = "black",
    face = "bold"
  )
)

p.covid

```

```{r covid_glm}
stargazer(zwalks.covid.glm, title="Regression Results", type = 'text') # nice print fo the glm binomial logistic regression
```

## Living and working zones

In the next step we take a look at the different different zones the counting stations are located in. The counting locations are distributed over 11 different zones types. These zones are defined by the state based on their usage and location. We summarized these zones types into five categories to simplify the analysis.

<!-- summarize zones -->

```{r zones}
# -------------------------- Zones --------------------------
### Bauzonen / usage and location zones
zstation.zones <- st_join(zstation.kreis, zzones,
                          join = st_within)

# look at the created zstation.zones
head(zstation.zones)
summary(zstation.zones)
unique((zstation.zones$typ))

# create a copy of typ
zstation.zones["zones"] <- zstation.zones$typ


### summarize zones
zstation.zones$zones[zstation.zones$zones == "W4"] <- "Living" # Viergeschossige  Wohnzone
zstation.zones$zones[zstation.zones$zones == "W5"] <- "Living" # Fünfgeschossige Wohnzone
zstation.zones$zones[zstation.zones$zones == "Z6"] <- "Center" # Sechsgeschossige Zentrumszone
zstation.zones$zones[zstation.zones$zones == "Z5"] <- "Center" # Fünfgeschossige Zentrumszone
zstation.zones$zones[zstation.zones$zones == "K"] <- "Center" # Kernzone
zstation.zones$zones[zstation.zones$zones == "F"] <- "Empty/Parking" # Freihaltezone
zstation.zones$zones[zstation.zones$zones == "FP"] <- "Empty/Parking" # Parking
zstation.zones$zones[zstation.zones$zones == "WLD"] <- "Recreation" # Wald
zstation.zones$zones[zstation.zones$zones == "QI"] <- "Recreation" # Quartier Erhaltungszone
zstation.zones$zones[zstation.zones$zones == "GWS"] <- "Recreation" # Gewässer
zstation.zones$zones[zstation.zones$zones == "IG I"] <- "Industry" # Industrie
```

<!--Zone plots -->

```{r tables_zones, results="markup"}
### table with zones, type and counting stations
table.zones.typ <- zstation.zones %>%
  group_by(typ, zones) %>%
  summarise(length(unique(abkuerzung))) %>%
  rename("Quantity of Stations" = "length(unique(abkuerzung))") %>%
  arrange(desc(zones))


# nice table on the counting stations per zone
knitr::kable(st_drop_geometry(table.zones.typ), #remove geometry
             align = "cr",
             caption = "Counting Stations per zone") %>%
  kable_styling(bootstrap_options = "striped") # from the kabelExtra package




### table with only the zones

table.zones <- zstation.zones %>%
  group_by(zones) %>%
  summarise(length(unique(abkuerzung))) %>%
  rename("Quantity of Stations" = "length(unique(abkuerzung))") %>%
  arrange(desc(zones))

#remove geometry

knitr::kable(st_drop_geometry(table.zones),
             align = "cr",
             caption = "Summarized Zones") %>%
  kable_styling(bootstrap_options = "striped") # from the kabelExtra package
```

<!-- Creatre one dataframe with all the information -->

```{r zwalks.difference}
# join everything into one table
zwalks.day.knr <- zwalks.day %>% left_join (dplyr::select(zstation.zones,
                                                fk_zaehler,
                                                knr,
                                                kname,
                                                zones),
                                            by = "fk_zaehler")



# group by-- week zones
zwalks.week.zones <- zwalks.day.knr %>%
  group_by(year, week, zones) %>%
  summarize(people_sum = sum(people_all))

# pivot wide table -- year zones
zwalks.week.zones <- zwalks.week.zones %>%
  pivot_wider(names_from = year,
              values_from = people_sum)

# calculate the difference between the 2020 and 2019
zwalks.week.zones["difference_people"] <- zwalks.week.zones['2020'] - zwalks.week.zones['2019']



#pivot wide table -- week knr
zwalks.week.knr <- zwalks.day.knr %>%
  group_by(year, week, knr) %>%
  summarize(people_sum = sum(people_all))


zwalks.week.knr <- zwalks.week.knr %>%
  pivot_wider(names_from = year,
              values_from = people_sum)

zwalks.week.knr["difference_people"] <- zwalks.week.knr['2020'] - zwalks.week.knr['2019'] 
```

As shown in the five plots below, the pandemic had a different impact on the different zones. Recreational zones and industrial zones counted higher numbers of people while Empty/Parking zones and Living zones where less frequently visited by people on foot or bicycle.

Except of the industry zones, the impact of the Lockdown is especially clear visible in March in all the zones. The lockdown effect seem to be vanishin in April and Mai.

```{r plot_zones, results="markup", fig.width = 12, fig.height = 6}
# plot: people per Zone

ggplot(zwalks.day.knr, aes(fill=year, x=factor(month, level=month.abb), y=people_all)) +
  geom_bar(position = "dodge", stat="identity") +
  ggtitle("Pedestrians and cyclists in the different zones") +
  xlab("Month") +
  ylab("Total People") +
  facet_wrap( ~ zones, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 45))+
  theme(legend.position = 'bottom')
```

## City Districts

In the last analysis part, we take a closer look at the twelve city districts. The twelve plots below show a comparison between 2020 and 2019 in the twelve districts. It seems that we are missing some measurements in district twelve for two months and district seven is missing completly as explained before.

Nevertheless, there are certain differences visible between the districts. District 1 and district 9 are hardest hit by the pandemic and the ongoing shift to home office. District 4, 8, and 12 seem to be more frequently visited by people on foot and bycicle during the pandemic. Our guess is that these districts have several free and open recreational areas. To test this statement, we would need additional information.

```{r plot_twelve_districts, results="markup", fig.width = 12, fig.height = 6}
# plot: people per district

ggplot(zwalks.day.knr, aes(fill=year, x=factor(month, level=month.abb), y=people_all)) +
  geom_bar(position = "dodge", stat="identity") +
  ggtitle("Pedestrians and cyclists in the districts ") +
  xlab("Month") +
  ylab("Total People") +
  facet_wrap( ~ knr, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45)) +
  labs(fill = "Year")+
  theme(legend.position = 'bottom')

```

In a second pair of plots, we only subtract the counted people in 2019 from the ones in 2020. Hence, the positive values indicate more counted people in 2020 and the negative show us when we counted more in 2019. Thhe numbers are presented on a weekly basis. We recommend to use both sets of plots to look at the differences between the two years to get the full picture. So seems that district 11 was heavily impacted by the pandemic, but when look at the number in the upper plot it becomes clear that there were just slightly less pedestrians and cyclists in the road throughout 2020.

# Video

In a last step, we decided to animate the difference in pedestrians and cyclists counted in 2020 compared to 2019 across time. For each month of 2020, we generate a color coded map. Dark green areas show that fewer pedestrians and bicyclists were counted compared to the previous year.In contrast, dark red areas show that more pedestrians and bicyclists where counted. This plot is combined with a second graph showing the number of COVID cases counted. If you look at zwalks_animation.avi, you can see how the COVID pandemic unfolds and how the number of pedestrians decreases as COVID cases rise.

```{r create_video, fig.show='hide'}
# -------------------------- animated video --------------------------

# create new directory & subdirectory
ifelse(!dir.exists(file.path("./Video/")), {dir.create(file.path("./Video/")); "Directory created"}, "Directory already exists")
ifelse(!dir.exists(file.path("./Video/png_output/")), {dir.create(file.path("./Video/png_output/")); "Subdirectory created"}, "Subirectory already exists")
png_output_path = "./Video/png_output/"


# -------------------------- prapare plot --------------------------
#load Stadtkreise
zkreis <- st_read("./Data/Stadtkreise/stzh.adm_stadtkreise_v.shp")
head(zkreis)
summary(zkreis)


#load countingstations
zstation <- st_read("./Data/Countingstation/taz.view_eco_standorte.shp")
head(zstation)
zstation <- cbind(zstation, st_coordinates(zstation))
summary(zstation)

# Basic colors
empty_kreis_color = "grey75"
missing_percentile_color = "grey75"
basic_color = "lightblue"
png_output_path = "./Video/png_output/"
zkreis$color <- basic_color


# -------------------------- prapare weekly data --------------------------

### calculate percentile
zwalks.week.knr <- zwalks.week.knr %>%
  group_by(knr) %>% 
  mutate(percrank=(ceiling(percent_rank(difference_people)*10)))
zwalks.week.knr

### Color Codes for percentiles
colfunc <- colorRampPalette(c("darkgreen", "red")) #ff9696
col_list <- colfunc(10)
plot(rep(1,10),col=col_list,pch=19,cex=3)
zwalks.week.knr$color = basic_color

## Set colors for percentile
for (i in 1:nrow(zwalks.week.knr)){
  current.percentile <- zwalks.week.knr$percrank[i]
  ### check if percentile is na
  if(is.na(current.percentile)){
    ### no percentile value found
    ### set color to missing_percentile_color
    zwalks.week.knr$color[i] <- missing_percentile_color
  } else {
    ### value is not na, so there should be a percentile
    ### make sure percentile is a number
    if(is.numeric(current.percentile)){
      if(current.percentile == 0){
        zwalks.week.knr$color[i] <- col_list[1]
      } else {
        ### assign color 
        zwalks.week.knr$color[i] <- col_list[current.percentile]
      }
    }
  }
}





### walk throguh weeks and generate plot image
tmp_current_date <- as.Date("2020-01-01")
for (i_week in 1:53){
  print(i_week)
  if(i_week ==1){
    tmp_current_date <- as.Date("2020-01-01")
  } else if (i_week == 2){
    tmp_current_date <- as.Date("2020-01-06")
  } else {
    tmp_current_date <- (tmp_current_date + 7)
  }
  print(tmp_current_date)
    
  ## bookpat
  print(i)
  print(paste("current week: " , toString(i_week)))


  zwalks.currentweek <- zwalks.week.knr %>% filter(week == i_week)
  ### debug
  print(zwalks.currentweek)
  ### assign colors to current week zkreis (zkreis= df to be plotted as a map)
  for (i_kreis in 1:nrow(zkreis)){
    ### reset color to base color in order to avoid missing values not being displayed as last week color
    zkreis$color[i_kreis] <- basic_color
    
    tmp.color <- zwalks.currentweek  %>% filter(knr == i_kreis)
    
    ### check if data for kreis exists.
    if(nrow(tmp.color) == 0) {
      ## kreis dosn't exists. Set color to missing_percentile_color 
      zkreis$color[i_kreis] <-missing_percentile_color
    } else {
      zkreis$color[i_kreis] <-tmp.color$color
    }
  }
  
  ### debug
  # print(zkreis)
  
  # -------------------------- plot map --------------------------
  
  plot.kreis <- ggplot(zkreis) +
    geom_sf(size = 1, color = "black", fill= zkreis$color) +
    geom_sf_text(aes(label = knr), colour = "black") +
    xlab(paste("Date: " , (tmp_current_date)))
  
  plot.station <- plot.kreis 
  
  
  ### Plot COVID Cases
  min_date_on_plot <- "2020-03-1"
  if(as.Date(tmp_current_date) > as.Date(min_date_on_plot)) {
    date_plot_end <- tmp_current_date
    print("date is larger")
  } else {
    date_plot_end <-toString(tmp_current_date)
    date_plot_end <- min_date_on_plot
    print("date is smaller")
  }
  
  # -------------------------- plot covid cases --------------------------
  ### Plot COVID Cases
  min_date_on_plot <- "2020-01-06"
  if(as.Date(tmp_current_date) > as.Date(min_date_on_plot)) {
    date_plot_end <- tmp_current_date
    print("date is larger")
  } else {
    date_plot_end <-toString(tmp_current_date)
    date_plot_end <- min_date_on_plot
    print("date is smaller")
  }
  
  
  covid.daterange<- filter(covid, between(covid$date, as.Date("2020-01-01"), as.Date(toString(date_plot_end))))
  #covid.daterange<- filter(covid, between(covid$date, as.Date("2020-01-01"), as.Date("2020-04-01")))
  print(covid.daterange$date)
  print(covid.daterange$new_cases_smoothed)
  
  p.covid <- ggplot(covid.daterange, aes(x=date, y=new_cases_smoothed)) +
    geom_line(color="green", size=0.7) + 
    ggtitle("COVID Cases") +
    ylab("") + 
    xlab("")
  #print(p.covid)
  
  ### Arrange multiple plots in one figure
  plot.arranged <- ggarrange(plot.station, p.covid , heights = c(2, 0.7),
                             ncol = 1, nrow = 2)
  print(plot.arranged )
  
  filenmae <- paste( png_output_path , '/', toString(i_week), "_week_zwalks.png" ,  sep = "" )
  print(filenmae)
  ggsave(filenmae)
}


# -------------------------- create video --------------------------


file_name<- list.files(path=png_output_path)
file_path <- paste(png_output_path , '/', file_name,   sep = "" )

file_list <- data.frame(file_name, file_path)
print(file_list)
file_list$sort_index = 0
print(file_list)



for (i in 1:nrow(file_list)){ 
  file_list$sort_index[i] <- strtoi(strsplit(file_list$file_name[i], "_")[[1]][1])
}

file_list_sorted <- arrange(file_list, sort_index)
file_list_sorted

av::av_encode_video(file_list_sorted$file_path, './Video/zurich_walks_video.avi', framerate = 2)


```

```{r vimeo_url}
suggest_embed("https://vimeo.com/516652769")
```

```{r vimeo_link, results="asis"}
embed_vimeo("516652769")
```

# Conclusion

With the mostly visual approach, we could show that the behavior of the people in Zurich changed compared to the previous year. While the impact of the lockdown on the counted people on the streets is no surprise, the clear increase during the summer months was rather unexpected.

The most outstanding part of our visual comparison was the increase of counted cyclists almost through the whole year.

This paper did primarily took a visually based comparison of the number of pedestrians and cyclists in 2019 and 2020 respectively during the two lockdowns in them. For further research, it would be necessary to quantify this identified impact statistically. A longer time series would be necessary to identify and address seasonality and trends in order to identify the impact of covid on the counted people.

\newpage

# References

## Data Sources

Roser, M., Ritchie, H., Ortiz-Ospina, E., & Hasell, J. (2020). Coronavirus Pandemic (COVID-19). Our World in Data. [\<https://ourworldindata.org/coronavirus/country/switzerland\>](https://ourworldindata.org/coronavirus/country/switzerland){.uri} (accessed 20.02.2021).

Stadt Zürich. (2021). Open Data Zürich. <https://data.stadt-zuerich.ch/> (accessed 10.02.2021).

Swiss Tourism Federation. (2021). Chronik Coronavirus. [\<https://www.stv-fst.ch/de/chronik-coronavirus\>](https://www.stv-fst.ch/de/chronik-coronavirus){.uri} (accessed 12.02.2021).

Widmer, P., & Robin, P. (2021). Zurich_walks_video. HSLU Bootcamp. <https://vimeo.com/516652769> (accessed 25.02.2021)

## Used R-Packages

```{r biblio_packages, include=FALSE}

used_packages <- names(sessionInfo()$otherPkgs)
knitr::write_bib(used_packages, file = './Sources/packages.bib')
```
