---
title: "Impact of Covid on Cyclists and Pedestrians in Zurich"
author: "Patrice.Robin, Patrik.Widmer"
date: "1 2 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE ,
                      message=FALSE,
                      warning = FALSE,
                      results = "hide"
                      )

```

# Introduction

Covid19 had a huge impact on our society. It changed the way we interact on a professional and social level. Home offices, lockdowns, and partially closed stores have had a particularly severe impact on urban centers.

In this paper, we will examine the impact of the pandemic on the number of pedestrians and bicycles in different areas of Zurich.

The city of Zurich automatically collects data on pedestrians and bicyclists through counting loops installed in the ground in different areas of the city. We can get a clear view of the impact of Covid19 on our daily lives by looking at the data of pedestrians and cyclists in 2020 and comparing it to the previous year.

```{r load packages}
# this line of code only runs, if these packages are required
if (!require(tidyverse)) {install.packages('"tidyverse"')}
if (!require(ggplot2)) {install.packages('ggplot2')}
if (!require(scales)) {install.packages('scales')}
if (!require(rgdal)) {install.packages('rgdal')}
if (!require(sf)) {install.packages('sf')}
if (!require(rasterVis)) {install.packages('rasterVis')}
if (!require(dplyr)) {install.packages('dplyr')}
if (!require(av)) {install.packages('av')}
if (!require(ggpubr)) {install.packages('ggpubr')}
```

```{r english_local}
#set local working directory to English
Sys.setlocale("LC_TIME", "English")
```

# Data
For our analysis, we relay on data from two sources:

-   All the data regarding Zurich are pulled from the official website of the city of Zurich *Open Data Stadt Zürich*

    -   Number of pedestrians and bicycles

    -   The location of the counting stations.

    -   The different *Stadtkreise* and their geodata points.

    -   The different districts in Zurich with the corresponding geo data points.

    -   A categorization of different areas based on their usage and building development.

-   The Covid19 data for Switzerland including the categories *cases, deaths, test, hospitalization, and vaccinations* are based on the research of Roser et al. (2020) and pulled from the OUr World in Data website.

```{r load_covid}
#load covid data
#source : https://ourworldindata.org/coronavirus/country/switzerland?country=~CHE
covid <- read.csv("covid-data.csv", encoding="UTF-8")
covid$date <- as.Date(covid$date,format = "%d.%m.%y")

```

```{r load_zwalks}
#load pedestrian and cyclist data
#source: https://data.stadt-zuerich.ch/dataset/ted_taz_verkehrszaehlungen_werte_fussgaenger_velo
zwalks2020 <- read.csv("2020_verkehrszaehlungen_werte_fussgaenger_velo.csv") #2019
zwalks2019 <- read.csv("2019_verkehrszaehlungen_werte_fussgaenger_velo.csv") #2020

zwalks <- dplyr::union(zwalks2020, zwalks2019)

```

```{r load_kreis}
#load Stadtkreise
#source https://data.stadt-zuerich.ch/dataset/geo_stadtkreise
zkreis <- st_read("./Stadtkreise/stzh.adm_stadtkreise_v.shp")

```

```{r load_zstation}
#load countingstations
#source https://data.stadt-zuerich.ch/dataset/geo_standorte_der_automatischen_fuss__und_vlozaehlungen

zstation <- st_read("./Countingstation/taz.view_eco_standorte.shp")
```

```{r load_zdistricts}
#Districts
#source: https://data.stadt-zuerich.ch/dataset/ktzh_quartieranalyse
zdistricts <- st_read("./Quartieranalyse/QUARTIERE_F.shp")
```

```{r load_zzones}
#Bauzonen
#sources https://data.stadt-zuerich.ch/dataset/geo_nutzungsplanung___kommunale_bau__und_zonenordnung__bzo__jahresendstand_2020
zzones <- st_read("./Bauzonen/afs.bzo_zone_v.shp")
```

## Preparation
In the initial dataset summarizes the counted pedestrians and cyclists into timeframes of 15 minutes. DUe to our interest into daily, weekly and monthly changes, we bring the counting data into needed format.

```{r clean_data}
#extract date from datetime
zwalks['date'] <- as.Date(zwalks$DATUM)

#lower cases column names
zwalks <- setNames(zwalks, tolower(names(zwalks)))

#replace NA with 0
zwalks[is.na(zwalks)] <- 0

#rename columns
zwalks <- zwalks %>%
  rename(datetime = datum)
```

```{r summarize_daily}
#summarize data on a daily basis
zwalks.day <-
  aggregate(zwalks[c("velo_in", "velo_out", "fuss_in", "fuss_out")],
            by = zwalks[c("fk_zaehler", "fk_standort", "date", "ost", "nord")],
            FUN = sum)

#summarize bicycles and pedestrians
zwalks.day['fuss_all'] <- zwalks.day$fuss_in + zwalks.day$fuss_out
zwalks.day['velo_all'] <- zwalks.day$velo_in + zwalks.day$velo_out
zwalks.day['people_all'] <-
  zwalks.day$velo_all + zwalks.day$fuss_all
```

```{r zwalks_date}
#add months and weekdays
zwalks.day$year <- format(zwalks.day$date, "%Y")
zwalks.day$month <- format(zwalks.day$date, "%B")
zwalks.day$week <- format(zwalks.day$date, "%W") #create weeks from date
zwalks.day$week <- as.numeric(zwalks.day$week) + 1 #make numeric and add 1 so that it goes from 1 to 53

#weekdays
zwalks.day$weekday <- weekdays(zwalks.day$date)

weekday_order <- c( "Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday","Sunday")
```

```{r join_zwalks_covid}
# left join zwalks agg and covid on date
zwalks.covid <- zwalks.agg %>% left_join(dplyr::select(covid,
                                                       date,
                                                       new_cases_smoothed),
                                         by = "date")
```

```{r agg_monthly}

zwalks.agg.month <-
  aggregate(zwalks.day[c("people_all", "fuss_all", "velo_all")],
            by = zwalks.day[c( "year", "month")],
            FUN = sum)
```

